---
title: "NFL Injuries"
author:
  - name: "Samantha Erne ^[Email: ernesm@miamioh.edu]"
    affiliation: Farmer School of Business, Miami University
  - name: "Brendan Beattie ^[Email: beattibs@miamioh.edu]"
    affiliation: Farmer School of Business, Miami University
  - name: "Jake Holroyd 2 ^[Email: holroyjm@miamioh.edu]"
    affiliation: Farmer School of Business, Miami University
  - name: "Peter Walsh ^[Email: walshpd2@miamioh.edu]"
    affiliation: Farmer School of Business, Miami University
  - name: "Dr. Fadel Megahed ^[Email: fmegahed@miamioh.edu]"
    affiliation: Farmer School of Business, Miami University
date: "2023-03-30"
output:
  html_document:
    code_folding: show
    code_download: TRUE
    number_sections: TRUE
    paged_df: TRUE
    toc: TRUE
    toc_float: TRUE
    theme: readable
---
NFL Injuries


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      progress = FALSE,
                      verbose = TRUE,
                      cache = TRUE)

```

## Required R Packages
```{r}
if(require(pacman)== FALSE) install.packages("pacman")
pacman::p_load(tidyverse, 
               tidycensus, # for getting the census data
               httr, jsonlite, # pkgs that we might use for API,
               janitor,
               lubridate,# for making a column name from row 1
               magrittr, rvest, nflreadr, nflfastR, reticulate, geosphere, stringr, devtools)
```



## Required Python Packages
```{r }
reticulate::py_install('bs4')
reticulate::py_install('pandas')
reticulate::py_install('lxml')
reticulate::py_install('requests')

```

---

## Importing NFL Schedule from 2019 to 2023 and Adding Home and Away Columns to the Dataframe

```{python complete_game_schedule_2019_23}
from bs4 import BeautifulSoup
import pandas as pd
import requests
import lxml.html as lh

# Loop through years 2019 to 2022
for year in range(2019, 2023):
    url_scores = f'https://www.pro-football-reference.com/years/{year}/games.htm'
    # Create object page
    page = requests.get(url_scores)
    doc = lh.fromstring(page.content)
    # Parse data that are stored between <tr>..</tr> of HTML
    tr_elements = doc.xpath('//table[@id="games"]/tbody/tr')

    # Extract required columns by their index
    required_columns = [0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13]

    # Create empty list to hold the rows of data
    data_rows = []

    # Process game data for each row
    for j in range(len(tr_elements)):
        T = tr_elements[j]
        # Check if row is a game row
        if len(T) == 14:
            # Iterate through each element of the row
            row_data = []
            for i in required_columns:
                data = T[i].text_content().strip()
                # Convert numerical values to integers
                if i in (7, 8, 9, 10, 11, 12, 13):
                    try:
                        data = int(data)
                    except:
                        pass
                # Append the data to the row list
                row_data.append(data)
            # Append the row to the data list
            data_rows.append(row_data)

    # Create a Pandas DataFrame from the processed data
    df_scores = pd.DataFrame(data_rows, columns=['week', 'day', 'date', 'time', 'winner/tie', 'at', 'loser/tie', 'boxscore', 'pts_w', 'pts_l', 'yds_w', 'tov_w', 'yds_l', 'tov_l'])

    # Add a season column with the year value
    df_scores['season'] = year

    # Append the DataFrame to a master DataFrame
    if year == 2019:
        df_master = df_scores
    else:
        df_master = pd.concat([df_master, df_scores], ignore_index=True)

def is_home_team(row):
  matchup = row["at"]
  winner_name = row['winner/tie']
  loser_name = row['loser/tie']
  if "@" in matchup:
    #home_team == loser_name
    home_team = loser_name
    return home_team == loser_name
  
  else:
    return False

# Apply the custom function to each row of the DataFrame
df_master["loser_is_home"] = df_master.apply(lambda row: is_home_team(row), axis=1)

df_master.drop(['at', 'boxscore'], axis = 1, inplace = True)
df_master = df_master[df_master["week"] != "Week"]
df_master = df_master[df_master["date"] != "Playoffs"]

## had to make changes to the "week" values that were strings: "WildCard", "Division", "ConfChamp", "SuperBowl"
df_master.to_csv("df_master_playoff_week_fix.csv", index = False)

from tkinter import Tk
from tkinter.filedialog import askopenfilename

# create a Tkinter root window
root = Tk()
root.withdraw()

# open a file dialog box to select the CSV file
file_path = askopenfilename(filetypes=[("CSV Files", "*.csv")])

# read the CSV file into a Pandas DataFrame
df_master_updated = pd.read_csv(file_path)

# close the root window
root.destroy()



## updated name of csv to make sure its not over written
#df_master_updated = pd.read_csv("df_master_playoff_week_fix_update.csv")


# Create empty lists for home and away teams
home_teams = []
away_teams = []

# Iterate over each row of the DataFrame
for index, row in df_master.iterrows():
    # Check if loser is home
    if row['loser_is_home'] == True:
        home_team = row['loser/tie']
        away_team = row['winner/tie']
    else:
        home_team = row['winner/tie']
        away_team = row['loser/tie']
    # Append home and away teams to the lists
    home_teams.append(home_team)
    away_teams.append(away_team)

# Add the new columns to the DataFrame
df_master_updated['Home_Team'] = home_teams
df_master_updated['Away_Team'] = away_teams

pd.set_option('display.max_columns', 50)
#df_master_updated
team_abbr = {
    'Arizona Cardinals': 'ARI',
    'Atlanta Falcons': 'ATL',
    'Baltimore Ravens': 'BAL',
    'Buffalo Bills': 'BUF',
    'Carolina Panthers': 'CAR',
    'Chicago Bears': 'CHI',
    'Cincinnati Bengals': 'CIN',
    'Cleveland Browns': 'CLE',
    'Dallas Cowboys': 'DAL',
    'Denver Broncos': 'DEN',
    'Detroit Lions': 'DET',
    'Green Bay Packers': 'GB',
    'Houston Texans': 'HOU',
    'Indianapolis Colts': 'IND',
    'Jacksonville Jaguars': 'JAX',
    'Kansas City Chiefs': 'KC',
    'Las Vegas Raiders': 'LV',
    'Los Angeles Chargers': 'LAC',
    'Los Angeles Rams': 'LAR',
    'Miami Dolphins': 'MIA',
    'Minnesota Vikings': 'MIN',
    'New England Patriots': 'NE',
    'New Orleans Saints': 'NO',
    'New York Giants': 'NYG',
    'New York Jets': 'NYJ',
    'Philadelphia Eagles': 'PHI',
    'Pittsburgh Steelers': 'PIT',
    'San Francisco 49ers': 'SF',
    'Seattle Seahawks': 'SEA',
    'Tampa Bay Buccaneers': 'TB',
    'Tennessee Titans': 'TEN',
    'Washington Football Team': 'WAS',
    'Washington Redskins': 'WAS',
    'Washington Commanders': 'WAS',
    'Oakland Raiders': 'OAK'
}
abbr_team = {v: k for k, v in team_abbr.items()}


def get_abbr(team_name):
    return team_abbr.get(team_name)

# Create new columns for home and away team abbreviations
df_master_updated['Home_abbr'] = df_master_updated.apply(lambda row: get_abbr(row['Home_Team']), axis=1)
df_master_updated['Away_abbr'] = df_master_updated.apply(lambda row: get_abbr(row['Away_Team']), axis=1)

# format the "week" column with leading zeros
df_master_updated['week'] = df_master_updated['week'].apply(lambda x: f'0{x}' if x < 10 else str(x))

# Create the new column
df_master_updated['game_id'] = df_master_updated['season'].astype(str) + '_' + df_master_updated['week'].astype(str).str.zfill(2) + '_' + df_master_updated['Away_abbr'] + '_' + df_master_updated['Home_abbr']

# Insert the new column at the beginning of the DataFrame
df_master_updated.insert(0, 'game_id', df_master_updated.pop('game_id'))
# Print the updated DataFrame


## for overwriting purposes
df_master_updated.to_csv('Master_Schedule_2019_23_updated_Rmd1.csv', index = False)
df_master_updated.head(20)
```


Created a `game_id` variable in order to join with the `injury_plays` table created using `nflfastR`. `game_id` formatted the same was as nflfastR encodes them is as follows: `season`_`week`_`away_abbr`_`home_abbr`. Finally, I had to change the playoff weeks that had names (e.g., WildCard, Division, ConfChamp, SuperBowl) manually because the NFL has changed the playoff format over the past few years.


## Joining the `master_schedule` with the `injury_plays` from the `nflfastR` package 

```{r pbp_injured_players_2019_23}

#setwd("Google Drive/Independent Study/final_files")
## this is equivalent to Master_Schedule_2019_23_updated_Rmd but did not want overwritten 
master_schedule = read_csv('Independent Study/final_files/Master_Schedule_2019_23_updated_Rmd1_update.csv')

start_season = 2019
end_season = 2022

pbp = nflfastR::load_pbp(start_season:end_season)

injury_plays = pbp %>% filter(grepl("injured", desc, ignore.case = TRUE))
injury_plays = injury_plays %>% unique()
injury_plays= injury_plays %>% select(game_id,home_team,away_team,season_type,week ,posteam,posteam_type,defteam,
                                      side_of_field,yardline_100,game_date,quarter_seconds_remaining,half_seconds_remaining,
                                      game_seconds_remaining,game_half,qtr,down, time,yrdln,ydstogo,desc)
injury_plays$injured_player = sub(".*(\\b\\w+\\.\\w+\\b) was injured.*", "\\1", injury_plays$desc)
injury_plays$injured_team = sub(".*(\\b\\w+-\\d+).*(\\b\\w+\\.\\w+\\b) was injured.*", "\\1", injury_plays$desc)


injury_plays = injury_plays %>% distinct()
#write_csv(injury_plays, "Rmd_files/fixing_injury_plays_manual.csv")

## manual changes made to names in order to make them properly merge later in the script
## issues with suffixes, hypenated names, etc.
injury_plays3 = read_csv("Independent Study/final_files/fixing_injury_plays_manual_update.csv")


combined_data = left_join(master_schedule, injury_plays3, by = "game_id")
## dropped 37 observations from the merge (3462 --> 3425)

combined_data %>% group_by(game_id) -> combined_grouped

counter_fun = function(x){
  ifelse(is.na(x), 0, length(x)) %>% unique()
}

combined_grouped %<>% 
  nest() %>%
  mutate(injured_players = map(.x = data, .f = magrittr::extract2, 'injured_player'),
         num_injuries = map_dbl(.x= injured_players, .f = counter_fun))

combined_grouped$injured_players <- ifelse(combined_grouped$injured_players == "NANA", "No Injury", combined_grouped$injured_players)
combined_grouped$injured_players <- ifelse(combined_grouped$injured_players == "NA", "No Injury", combined_grouped$injured_players)

## separates the names to only include player names that got injured during the game
combined_grouped$injured_players <- sapply(combined_grouped$injured_players, function(x) ifelse(is.character(x), paste(sapply(strsplit(x, ", "), function(y) trimws(y)), collapse = ", "), x))

combined_grouped = combined_grouped[,-2]

schedule_injury_2019_23_complete = left_join(master_schedule, combined_grouped, by = "game_id")
#write_csv(schedule_injury_2019_23_complete, "schedule_injury_2019_23_complete.csv")
## change weeks variable to have playoffs encoded as numbers instead of "WildCard", "Division", "ConfChamp", "SuperBowl

new = read_csv("Independent Study/final_files/schedule_injury_2019_23_complete_numeric_weeks.csv")

## Adding the number of snaps that took place in a game

play_counts = as.data.frame(table(pbp$game_id))
names(play_counts) = c("game_id", "num_plays")
#play_counts <- play_counts[-c(267, 536, 821, 1105), ]
## Reset index
row.names(play_counts) <- NULL
play_counts <- data.frame(play_counts)

## Main Issue: one data set has the Los Angeles Rams abbreviation as LA and the other dataset has the abbreviation of LAR (71 observations)
# Some lines were not properly reading (duplicate games with all NAs in one duplicate besides the num_plays and other duplicate has all the information and NAs for num_play)

#write_csv(play_counts, "play_counts_LA_to_LAR.csv")

play_counts4 = read_csv("Independent Study/final_files/play_counts_LA_to_LAR_update.csv")
### May not need this line, causing problems in later sections
#play_counts$game_id <- gsub("(^|_)LA(_|$)", "\\1LAR\\2", play_counts$game_id)


schedule_injury_2019_23_complete_combined = merge(new, play_counts4, by = "game_id")
#rows_dropped = anti_join(schedule_injury_2019_23_complete, play_counts4, by = "game_id")
## two rows dropped after this merge

## Manually changing some of the injury_players because some of them are not formatted in a proper way (i.e., hyphens and spaces in name, different formatting than the rest, etc.)
#write_csv(schedule_injury_2019_23_complete_combined, 'schedule_injury_2019_23_complete_upated.csv')

```


## Adding Stadium Name, Surface, Surface_Type, and whether or not the stadium has a dome

```{r adding_surface_stadium_name_and_dome}
## this is equivalent of schedule_injury_2019_23_complete_upated but made new copy to make sure the manual changes to names was not overwritten
schedule_injury_2019_23_complete_updated = read_csv('Independent Study/final_files/schedule_injury_2019_23_complete_upated_Rmd.csv')
## changing LAR to LA to match the rest of the dataset
#schedule_injury_2019_23_complete_updated$game_id <- gsub("LAR", "LA", schedule_injury_2019_23_complete_updated$game_id)


# Create a list of dictionaries that maps team abbreviations to the surface, stadium, dome, and surface_type they play on
team_surface <- list(
  ARI = list(surface = "Bermuda Grass", surface_type = "Grass", stadium = "State Farm Stadium", dome = TRUE),
  ATL = list(surface = "FieldTurf Revolution", surface_type = "Turf", stadium = "Mercedes-Benz Stadium", dome = TRUE),
  BAL = list(surface = "Bermuda Grass", surface_type = "Grass", stadium = "M&T Bank Stadium", dome = FALSE),
  BUF = list(surface = "A-Turf Titan", surface_type = "Turf", stadium = "Highmark Stadium", dome = FALSE),
  CAR = list(surface = "Bermuda Grass", surface_type = "Grass", stadium = "Bank of America Stadium", dome = FALSE),
  CHI = list(surface = "Kentucky Bluegrass", surface_type = "Grass", stadium = "Soldier Field", dome = FALSE),
  CIN = list(surface = "UBU Sports Speed S5-M Synthetic Turf", surface_type = "Turf", stadium = "Paul Brown Stadium", dome = FALSE),
  CLE = list(surface = "Kentucky Bluegrass", surface_type = "Grass", stadium = "FirstEnergy Stadium", dome = FALSE),
  DAL = list(surface = "Hellas Matrix Turf", surface_type = "Turf", stadium = "AT&T Stadium", dome = TRUE),
  DEN = list(surface = "Kentucky Bluegrass", surface_type = "Grass", stadium = "Empower Field at Mile High", dome = FALSE),
  DET = list(surface = "FieldTurf Classic HD", surface_type = "Turf", stadium = "Ford Field", dome = TRUE),
  GB = list(surface = "Desso GrassMaster", surface_type = "Hybrid", stadium = "Lambeau Field", dome = FALSE),
  HOU = list(surface = "Hellas Matrix Turf", surface_type = "Turf", stadium = "NRG Stadium", dome = TRUE),
  IND = list(surface = "Shaw Sports Momentum Pro", surface_type = "Turf", stadium = "Lucas Oil Stadium", dome = TRUE),
  JAX = list(surface = "Bermuda Grass", surface_type = "Grass", stadium = "TIAA Bank Field", dome = FALSE),
  KC = list(surface = "Bermuda Grass", surface_type = "Grass", stadium = "Arrowhead Stadium", dome = FALSE),
  LV = list(surface = "Bermuda Grass", surface_type = "Grass", stadium = "Allegiant Stadium", dome = TRUE),
  LAC = list(surface = "Hellas Matrix Turf", surface_type = "Turf", stadium = "SoFi Stadium", dome = TRUE),
  LAR = list(surface = "Hellas Matrix Turf", surface_type = "Turf", stadium = "SoFi Stadium", dome = TRUE),
  MIA = list(surface = "Bermuda Grass", surface_type = "Grass", stadium = "Hard Rock Stadium", dome = TRUE),
  MIN = list(surface = "UBU Sports Speed S5-M", surface_type = "Turf", stadium = "U.S. Bank Stadium", dome = TRUE),
  NE = list(surface = "FieldTurf CORE", surface_type = "Turf", stadium = "Gillette Stadium", dome = FALSE),
  NO = list(surface = "FieldTurf Revolution 360", surface_type = "Turf", stadium = "Caesars Superdome", dome = TRUE),
  NYG = list(surface = "FieldTurf Classic HD", surface_type = "Turf", stadium = "MetLife Stadium", dome = FALSE),
  NYJ = list(surface = "FieldTurf Classic HD", surface_type = "Turf", stadium = "MetLife Stadium", dome = FALSE),
  PHI = list(surface = "Desso GrassMaster", surface_type = "Hybrid", stadium = "Lincoln Financial Field", dome = FALSE),
  PIT = list(surface = "Kentucky Bluegrass", surface_type = "Grass", stadium = "Heinz Field", dome = FALSE),
  SF = list(surface = "Bermuda Grass,Perennial Ryegrass Mixture", surface_type = "Grass", stadium = "Levi's Stadium", dome = FALSE),
  SEA = list(surface = "FieldTurf Revolution 360", surface_type = "Turf", stadium = "Lumen Field", dome = FALSE),
  TB = list(surface = "Bermuda Grass", surface_type = "Grass", stadium = "Raymond James Stadium", dome = FALSE),
  TEN = list(surface = "Bermuda Grass", surface_type = "Grass", stadium = "Nissan Stadium", dome = FALSE),
  WAS = list(surface = "Bermuda Grass", surface_type = "Grass", stadium = "FedExField", dome = FALSE),
  OAK = list(surface = "Bermuda Grass", surface_type = "Grass", stadium = "Oakland Coliseum", dome = FALSE)
)

team_surface_df <- map_df(team_surface, ~.x %>% as_tibble() %>% set_names(c("surface", "surface_type", "stadium", "dome")), .id = "team")


team_surface_df <- team_surface_df %>%
  rename(Home_abbr = team)

library(dplyr)

#schedule_injury_2019_23_complete_updated <- schedule_injury_2019_23_complete_updated %>% select(-Home_abbr, everything(), Home_abbr)

#schedule_injury_2019_23_complete_updated$Home_abbr <- ifelse(schedule_injury_2019_23_complete_updated$Home_abbr == "LAR", "LA", schedule_injury_2019_23_complete_updated$Home_abbr)
#schedule_injury_2019_23_complete_updated$Away_abbr <- ifelse(schedule_injury_2019_23_complete_updated$Away_abbr == "LAR", "LA", schedule_injury_2019_23_complete_updated$Away_abbr)


df_master_schedule_injury_surface_2019_23 <- schedule_injury_2019_23_complete_updated %>%
  left_join(team_surface_df, by = "Home_abbr")

colSums(is.na(df_master_schedule_injury_surface_2019_23))
## checking to make sure there are no NAs present after merge is complete
#write_csv(df_master_schedule_injury_surface_2019_23, 'df_master_schedule_injury_surface_2019_23_Rmd.csv')
head(df_master_schedule_injury_surface_2019_23,20)


```

## Adding Weather Script to `df_master_schedule_injury_surface_2019_23`

```{r adding_weather_to_df_master_above}
pacman::p_load(tidyverse, rvest, stringr)

## this is the equivalent of df_master_schedule_injury_surface_2019_23_Rmd.csv but made copy to make sure it was not overwritten during knitting process

## changing playoff weeks from numbers to "Wildcard", "Division", etc.
schedule = read.csv('Independent Study/final_files/df_master_schedule_injury_surface_2019_23_Rmd_no_weather.csv')
schedule$date = schedule$date %>% as.Date(format = "%m/%d/%y")

results <- list()

# loop through each row of the data frame
for (i in 1:nrow(schedule)) {
  # extract the variables for the current row
  season = schedule$season[[i]]
  week = schedule$week[[i]]
  date = schedule$date[[i]]
  away = tail(strsplit(schedule$Away_Team[i], " ")[[1]], 1)
  away = ifelse(away == 'Team' & date >= '2020-09-13' & date <= '2020-09-20', 
                'redskins',
                ifelse(away == 'Team' & date > '2020-09-20' & date <= '2020-11-22', 
                       'football%20team',
                       ifelse(away == 'Team' & date > '2020-11-22' & date <= '2020-12-13',
                              'Washington',
                              ifelse(away == 'Team' & date == '2020-12-20',
                                     'football%20team',
                                     ifelse((away == 'Team' | away == 'Commanders') & date > '2020-12-20',
                                            'Washington',
                                            tail(strsplit(schedule$Away_Team[i], " ")[[1]], 1))))))
  
  home = tail(strsplit(schedule$Home_Team[i], " ")[[1]], 1)
  home = ifelse(home == 'Team' & date >= '2020-09-13' & date <= '2020-09-20', 
                'redskins',
                ifelse(home == 'Team' & date > '2020-09-20' & date <= '2020-11-22', 
                       'football%20team',
                       ifelse(home == 'Team' & date > '2020-11-22' & date <= '2020-12-13',
                              'washington',
                              ifelse(home == 'Team' & date == '2020-12-20',
                                     'football%20team',
                                     ifelse((home == 'Team' | home == 'Commanders') & date > '2020-12-20',
                                            'washington',
                                            tail(strsplit(schedule$Home_Team[i], " ")[[1]], 1))))))
  
  
  # construct the URL and read the HTML
  url = ifelse(week == 'WildCard', 
               paste0('https://www.nflweather.com/game/', season, '/',  week,'-weekend/', away, '-at-', home),
               ifelse(week == 'Division', paste0('https://www.nflweather.com/game/', season, '/',  week,'al-playoffs/', away, '-at-', home),
                      ifelse(week == 'ConfChamp', paste0('https://www.nflweather.com/game/', season, '/%20conf-championships/', away, '-at-', home),
                             ifelse(week == 'SuperBowl' & season == '2021', paste0('https://www.nflweather.com/game/', season, '/',  week, '/', home, '-at-', away),
                                    ifelse(week == 'SuperBowl', paste0('https://www.nflweather.com/game/', season, '/', week, '/', away, '-at-', home),
                                           paste0('https://www.nflweather.com/game/', season, '/week-',  week, '/', away, '-at-', home))))))
  
  read_url <- read_html(url)
  
  # extract the weather data from each quarter
  q1_temp <- html_text2(html_elements(read_url, 'body > div > div:nth-child(4) > div > div.span10 > div:nth-child(7) > div:nth-child(1) > p:nth-child(5) > b')) |> 
    str_extract("\\d+") |> 
    as.numeric()
  q2_temp <- html_text2(html_elements(read_url, 'body > div > div:nth-child(4) > div > div.span10 > div:nth-child(7) > div:nth-child(2) > p:nth-child(5) > b')) |> 
    str_extract("\\d+") |> 
    as.numeric()
  q3_temp <- html_text2(html_elements(read_url, 'body > div > div:nth-child(4) > div > div.span10 > div:nth-child(7) > div:nth-child(3) > p:nth-child(5) > b')) |> 
    str_extract("\\d+") |> 
    as.numeric()
  q4_temp <- html_text2(html_elements(read_url, 'body > div > div:nth-child(4) > div > div.span10 > div:nth-child(7) > div:nth-child(4) > p:nth-child(5) > b')) |> 
    str_extract("\\d+") |> 
    as.numeric()
  
  q1_feels_like = html_text2(html_elements(read_url, 'body > div > div:nth-child(4) > div > div.span10 > div:nth-child(7) > div:nth-child(1) > p:nth-child(6) > b')) |> 
    str_extract("\\d+") |> 
    as.numeric()
  q2_feels_like = html_text2(html_elements(read_url, 'body > div > div:nth-child(4) > div > div.span10 > div:nth-child(7) > div:nth-child(2) > p:nth-child(6) > b')) |> 
    str_extract("\\d+") |> 
    as.numeric()
  q3_feels_like = html_text2(html_elements(read_url, 'body > div > div:nth-child(4) > div > div.span10 > div:nth-child(7) > div:nth-child(3) > p:nth-child(6) > b')) |> 
    str_extract("\\d+") |> 
    as.numeric()
  q4_feels_like = html_text2(html_elements(read_url, 'body > div > div:nth-child(4) > div > div.span10 > div:nth-child(7) > div:nth-child(4) > p:nth-child(6) > b')) |> 
    str_extract("\\d+") |> 
    as.numeric()
  
  
  q1_wind = html_text2(html_elements(read_url, 'body > div > div:nth-child(4) > div > div.span10 > div:nth-child(7) > div:nth-child(1) > p:nth-child(7) > b')) |> 
    str_extract("\\d+") |> 
    as.numeric()
  q2_wind = html_text2(html_elements(read_url, 'body > div > div:nth-child(4) > div > div.span10 > div:nth-child(7) > div:nth-child(2) > p:nth-child(7) > b')) |> 
    str_extract("\\d+") |> 
    as.numeric()
  q3_wind = html_text2(html_elements(read_url, 'body > div > div:nth-child(4) > div > div.span10 > div:nth-child(7) > div:nth-child(3) > p:nth-child(7) > b')) |> 
    str_extract("\\d+") |> 
    as.numeric()
  q4_wind = html_text2(html_elements(read_url, 'body > div > div:nth-child(4) > div > div.span10 > div:nth-child(7) > div:nth-child(4) > p:nth-child(7) > b')) |> 
    str_extract("\\d+") |> 
    as.numeric()
  
  q1_humidity = html_text2(html_elements(read_url, 'body > div > div:nth-child(4) > div > div.span10 > div:nth-child(7) > div:nth-child(1) > p:nth-child(8) > b')) |> 
    str_extract("\\d+") |> 
    as.numeric()
  q2_humidity = html_text2(html_elements(read_url, 'body > div > div:nth-child(4) > div > div.span10 > div:nth-child(7) > div:nth-child(2) > p:nth-child(8) > b')) |> 
    str_extract("\\d+") |> 
    as.numeric()
  q3_humidity = html_text2(html_elements(read_url, 'body > div > div:nth-child(4) > div > div.span10 > div:nth-child(7) > div:nth-child(3) > p:nth-child(8) > b')) |> 
    str_extract("\\d+") |> 
    as.numeric()
  q4_humidity = html_text2(html_elements(read_url, 'body > div > div:nth-child(4) > div > div.span10 > div:nth-child(7) > div:nth-child(4) > p:nth-child(8) > b')) |> 
    str_extract("\\d+") |> 
    as.numeric()
  
  q1_percipitation_prob = html_text2(html_elements(read_url, 'body > div > div:nth-child(4) > div > div.span10 > div:nth-child(7) > div:nth-child(1) > p:nth-child(13) > b')) |> 
    str_extract("\\d+") |> 
    as.numeric()
  q2_percipitation_prob = html_text2(html_elements(read_url, 'body > div > div:nth-child(4) > div > div.span10 > div:nth-child(7) > div:nth-child(2) > p:nth-child(13) > b')) |> 
    str_extract("\\d+") |> 
    as.numeric()
  q3_percipitation_prob = html_text2(html_elements(read_url, 'body > div > div:nth-child(4) > div > div.span10 > div:nth-child(7) > div:nth-child(3) > p:nth-child(13) > b')) |> 
    str_extract("\\d+") |> 
    as.numeric()
  q4_percipitation_prob = html_text2(html_elements(read_url, 'body > div > div:nth-child(4) > div > div.span10 > div:nth-child(7) > div:nth-child(4) > p:nth-child(13) > b')) |> 
    str_extract("\\d+") |> 
    as.numeric()
  
  # calculate the average weather data for the game
  avg_temp = mean(c(q1_temp, q2_temp, q3_temp, q4_temp))
  avg_feels_like = mean(c(q1_feels_like, q2_feels_like, q3_feels_like, q4_feels_like))
  avg_wind_mph = mean(c(q1_wind, q2_wind, q3_wind, q4_wind))
  avg_humidity_percent = mean(c(q1_humidity, q2_humidity, q3_humidity, q4_humidity))
  avg_percipitation_prob_percent = mean(c(q1_percipitation_prob, q2_percipitation_prob, q3_percipitation_prob, q4_percipitation_prob))
  
  
  
  # create a data frame with the results for the current row
  result <- data.frame(week, season, date, away, home, avg_temp, avg_feels_like, avg_wind_mph, avg_humidity_percent, avg_percipitation_prob_percent)
  
  # add the data frame to the results list
  results[[i]] = result
}

# combine the results into a single data frame
weather = do.call(rbind, results)

#write.csv(weather, "Game_Weather_2019-23.csv")

schedule = schedule |> 
  dplyr::mutate(
    Avg_Temp = weather$avg_temp,
    Avg_Feels_Like = weather$avg_feels_like,
    Avg_Wind_MPH = weather$avg_wind_mph,
    Avg_Humidity_Percent = weather$avg_humidity_percent,
    Avg_Percipitation_Prob_Percent = weather$avg_percipitation_prob_percent
  )

#write.csv(schedule, "df_master_schedule_injury_surface_2019_23_weather_Rmd.csv")
## fix the last SuperBowl Info
head(schedule, 20)

```

## Adding Latitude & Longitude based on Distance from Away Team's Stadium to the Home Team's Stadium (distance traveled for away games)

```{r}
library(purrr)
df_master_schedule_injury_surface_2019_23_weather = read_csv("Independent Study/final_files/df_master_schedule_injury_surface_2019_23_weather_Rmd_update.csv")
df_master_schedule_injury_surface_2019_23_weather = df_master_schedule_injury_surface_2019_23_weather[,-1]

## changing LAR to LA in the game_id
#df_master_schedule_injury_surface_2019_23_weather$game_id <- gsub("LAR", "LA", df_master_schedule_injury_surface_2019_23_weather$game_id)


#df_master_schedule_injury_surface_2019_23_weather$Away_abbr <- ifelse(df_master_schedule_injury_surface_2019_23_weather$Away_abbr == "LAR", "LA", df_master_schedule_injury_surface_2019_23_weather$Away_abbr)

#df_master_schedule_injury_surface_2019_23_weather$Home_abbr <- ifelse(df_master_schedule_injury_surface_2019_23_weather$Home_abbr == "LAR", "LA", df_master_schedule_injury_surface_2019_23_weather$Home_abbr)

colSums(is.na(df_master_schedule_injury_surface_2019_23_weather))


nfl_coordinates = read_csv("Independent Study/final_files/NFL_Latitude_Longitude_trial.csv")

nfl_coordinates <- nfl_coordinates %>%
  mutate(team_abbr = case_when(
    Team == "Arizona Cardinals" ~ "ARI",
    Team == "Atlanta Falcons" ~ "ATL",
    Team == "Baltimore Ravens" ~ "BAL",
    Team == "Buffalo Bills" ~ "BUF",
    Team == "Carolina Panthers" ~ "CAR",
    Team == "Chicago Bears" ~ "CHI",
    Team == "Cincinnati Bengals" ~ "CIN",
    Team == "Cleveland Browns" ~ "CLE",
    Team == "Dallas Cowboys" ~ "DAL",
    Team == "Denver Broncos" ~ "DEN",
    Team == "Detroit Lions" ~ "DET",
    Team == "Green Bay Packers" ~ "GB",
    Team == "Houston Texans" ~ "HOU",
    Team == "Indianapolis Colts" ~ "IND",
    Team == "Jacksonville Jaguars" ~ "JAX",
    Team == "Kansas City Chiefs" ~ "KC",
    Team == "Las Vegas Raiders" ~ "LV",
    Team == "Los Angeles Chargers" ~ "LAC",
    Team == "Los Angeles Rams" ~ "LAR",
    Team == "Miami Dolphins" ~ "MIA",
    Team == "Minnesota Vikings" ~ "MIN",
    Team == "New England Patriots" ~ "NE",
    Team == "New Orleans Saints" ~ "NO",
    Team == "New York Giants" ~ "NYG",
    Team == "New York Jets" ~ "NYJ",
    Team == "Philadelphia Eagles" ~ "PHI",
    Team == "Pittsburgh Steelers" ~ "PIT",
    Team == "San Francisco 49ers" ~ "SF",
    Team == "Seattle Seahawks" ~ "SEA",
    Team == "Tampa Bay Buccaneers" ~ "TB",
    Team == "Tennessee Titans" ~ "TEN",
    Team == "Washington Redskins" | 
    Team == "Washington Football Team" | 
    Team == "Washington Commanders" ~ "WAS",
    Team == "Oakland Raiders" ~ "OAK",
    TRUE ~ NA_character_
  ))

library(dplyr)
duplicated(nfl_coordinates$team_abbr)

# Create a lookup table for team abbreviations, latitude, and longitude
team_lookup <- nfl_coordinates %>% 
  select(team_abbr, Latitude, Longitude)

# Join the team_lookup to the master schedule
df_master_surface_coords <- df_master_schedule_injury_surface_2019_23_weather %>%
  left_join(team_lookup, by = c("Home_abbr" = "team_abbr")) %>%
  rename(Home_latitude = Latitude, Home_longitude = Longitude) %>%
  left_join(team_lookup, by = c("Away_abbr" = "team_abbr")) %>%
  rename(Away_latitude = Latitude, Away_longitude = Longitude)


colSums(is.na(df_master_surface_coords))
library(geosphere)

haversine_distance <- function(lat1, lon1, lat2, lon2) {
  # Convert latitudes and longitudes from degrees to radians
  lat1_rad <- lat1 * (pi / 180)
  lon1_rad <- lon1 * (pi / 180)
  lat2_rad <- lat2 * (pi / 180)
  lon2_rad <- lon2 * (pi / 180)

  # Haversine formula
  delta_lat <- lat2_rad - lat1_rad
  delta_lon <- lon2_rad - lon1_rad

  a <- sin(delta_lat / 2)^2 + cos(lat1_rad) * cos(lat2_rad) * sin(delta_lon / 2)^2
  c <- 2 * asin(sqrt(a))

  # Earth's radius in miles (approximately 3,958.8 miles)
  earth_radius <- 3958.8

  # Calculate the distance
  distance <- earth_radius * c

  return(distance)
}
Home_latitude <- round(df_master_surface_coords[["Home_latitude"]],4)
Home_longitude <- round(df_master_surface_coords[["Home_longitude"]],4)
Away_latitude <- round(df_master_surface_coords[["Away_latitude"]],4)
Away_longitude <- round(df_master_surface_coords[["Away_longitude"]],4)


df_master_surface_coords$distance_miles <- haversine_distance(Home_latitude, Home_longitude, Away_latitude, Away_longitude)

#df_master_surface_coords$distance_miles <- distHaversine(
  #p1 = df_master_surface_coords[, c("Home_longitude", "Home_latitude")],
  #p2 = df_master_surface_coords[, c("Away_longitude", "Away_latitude")],
  #r = 3959  # Earth radius in miles
#)
#df_master_surface_coords = df_master_surface_coords[,-1]

#write.csv(df_master_surface_coords, "Rmd_files/df_master_schedule_injury_surface_2019_23_weather_distance2.csv")
## STILL NEED TO CHANGE THE GAMES THAT WERE INTERNATIONAL OR PLAYED AT NEUTRAL SITES (SUPERBOWL)

## ADD BACK THE TWO SUPERBOWL GAMES THAT WERE DROPPED AND ASSOCIATED INJURIES (2019 AND 2020 SB)
head(df_master_surface_coords,20)
#write_csv(df_master_surface_coords,"df_master_schedule_injury_surface_2019_23_weather_distance_update.csv")


```


## Adding Days since both teams played their last game

```{r days_since_last_game_merge}

library(dplyr)

# read in the tables
df_master_surface_coords <- read_csv("Independent Study/final_files/df_master_schedule_injury_surface_2019_23_weather_distance_update.csv")
NFL_Days_Since_Last_Game <- read_csv("Independent Study/final_files/NFL_Days_Since_Last_Game.csv")


NFL_Days_Since_Last_Game$date = NFL_Days_Since_Last_Game$Date %>% as.Date(format = "%m/%d/%y")

# join the tables
df_master_surface_coords <- df_master_surface_coords %>%
  # join with NFL_Days_Since_Last_Game by home_team, away_team and date_of_game
  left_join(NFL_Days_Since_Last_Game, by = c("Home_Team" = "Home", "Away_Team" = "Away", "date" = "date"))
colSums(is.na(df_master_surface_coords))
#df_master_surface_coords = df_master_surface_coords[,-1]
df_master_surface_coords = df_master_surface_coords[,-37]
df_master_surface_coords = df_master_surface_coords[,-37]

head(df_master_surface_coords,20)
#write_csv(df_master_surface_coords, "df_master_schedule_injury_surface_2019_23_weather_distance_days_since_last_game.csv")

```

```{r players_missing_games_from_injury}
# Define the URL template
url_template <- "https://www.prosportstransactions.com/football/Search/SearchResults.php?Player=&Team=&BeginDate=2019-09-08&EndDate=2023-01-29&InjuriesChkBx=yes&submit=Search&start=%d"

# Specify the starting value for the start parameter
start_value <- 0

# Initialize an empty data.frame to store the data
games_missed_from_injury <- data.frame()

# Repeat the loop until you have scraped all pages (105 pages)
while (start_value <= 2600) {
  # Define the URL
  url <- sprintf(url_template, start_value)
  
  # Read the HTML content from the URL
  html_content <- read_html(url)
  
  # Extract the table from the HTML content
  table_html <- html_content %>%
    html_node("table") %>%
    html_table()
  
  # Check if the table is empty (no more pages)
  if (nrow(table_html) == 0) {
    break
  }
  
  # Specify the column names
  colnames(table_html) <- c("Date", "Team","Acquired" ,"Relinquished", "Notes")
  
  # Store the table data in a data.frame
  df <- as.data.frame(table_html)
  
  # Append the data to the all-data data.frame
  games_missed_from_injury <- rbind(games_missed_from_injury, df)
  
  # Update the start value to scrape the next page
  start_value <- start_value + 25
}

## removing additional rows with no information
games_missed_from_injury = subset(games_missed_from_injury, Date != "Date")
## removing Acquired variable because it contains no information
games_missed_from_injury = subset(games_missed_from_injury, select = -Acquired)

library(stringr)

## removing the bullet in front of the plays name who was injured 
games_missed_from_injury$Relinquished <- str_remove(games_missed_from_injury$Relinquished, "• ")
## removing parenthesis after name (usually containing the players full name or their birthdate) to ensure that all names are formatted the same way
games_missed_from_injury$Relinquished <- sub("\\(.*\\)", "", games_missed_from_injury$Relinquished)
## removing the slashes after the name of the player who was injury (usually contains the players nickname(s)) to ensure that all names are formatted the same way
games_missed_from_injury$Relinquished <- sapply(strsplit(games_missed_from_injury$Relinquished, "/"), "[", 1)
## formatting 'Date' variable as a date in order to count the number of days since last game was played
games_missed_from_injury$Date <- as.Date(games_missed_from_injury$Date)

## adding year column in order to create unique id to merge tables
#games_missed_from_injury$Year = year(games_missed_from_injury$Date)


## splitting the relinquished column to have a first_name and last_name column
games_missed_from_injury$first_name <- gsub("^(\\S+)\\s.*$", "\\1", games_missed_from_injury$Relinquished)
games_missed_from_injury$last_name <- gsub("^\\S+\\s(.*)$", "\\1", games_missed_from_injury$Relinquished)

## removing suffixes to have the data structured the same way as those from fast_scraper_roster function within nflfastR
#games_missed_from_injury$last_name <- gsub("\\s.*", "", games_missed_from_injury$last_name)
games_missed_from_injury$last_name <- sapply(games_missed_from_injury$last_name, function(x) {
  if (grepl(" ", x)) {
    parts <- strsplit(x, " ")[[1]]
    if (parts[length(parts)] %in% c("Jr.", "Sr.","II" ,"III", "IV", "V")) {
      return(parts[length(parts) - 1])
    } else {
      return(parts[length(parts)])
    }
  } else {
    return(x)
  }
})
## adding season variable to merge with the table below (accounting for games played in the first few weeks of the new year as included in the previous season)
# Create a "season" variable based on the "Date" column
games_missed_from_injury$season <- format(games_missed_from_injury$Date, "%Y")

# Adjust the season values for the 2019 season
#games_missed_from_injury$season[games_missed_from_injury$Date < as.Date("2019-08-08")] <- as.character(as.numeric(format(games_missed_from_injury$Date[games_missed_from_injury$Date < as.Date("2019-08-08")], "%Y")) - 1)

# Adjust the season values for the 2020 season
games_missed_from_injury$season[games_missed_from_injury$Date >= as.Date("2019-06-27") & games_missed_from_injury$Date <= as.Date("2020-02-28")] <- "2019"

# Adjust the season values for the 2021 season
games_missed_from_injury$season[games_missed_from_injury$Date > as.Date("2020-05-08") & games_missed_from_injury$Date < as.Date("2021-02-28")] <- "2020"

# Adjust the season values for the 2021 season
games_missed_from_injury$season[games_missed_from_injury$Date > as.Date("2021-05-08") & games_missed_from_injury$Date < as.Date("2022-03-10")] <- "2021"

# Adjust the season values for the 2022 season
games_missed_from_injury$season[games_missed_from_injury$Date >= as.Date("2022-03-11")] <- "2022"


## adding team abbreviations in order to create a unique id to join the player_position table
nfl_teams <- data.frame(
  team_name = c("Cardinals", "Falcons", "Ravens", "Bills", "Panthers", "Bears", "Bengals", "Browns", "Cowboys", "Broncos", "Lions", "Packers", 
                "Texans", "Colts", "Jaguars", "Chiefs", "Raiders", "Chargers", "Rams", "Dolphins", "Vikings", "Patriots", "Saints", "Giants", 
                "Jets", "Eagles", "Steelers", "49ers", "Seahawks", "Buccaneers", "Titans", "Redskins", "Washington", "Commanders"),
  team_abbreviation = c("ARI", "ATL", "BAL", "BUF", "CAR", "CHI", "CIN", "CLE", "DAL", "DEN", "DET", "GB", "HOU", "IND", "JAX", "KC", 
                        "LV", "LAC", "LAR", "MIA", "MIN", "NE", "NO", "NYG", "NYJ", "PHI", "PIT", "SF", "SEA", "TB", "TEN", "WAS", "WAS", "WAS")
)

# Join the nfl_teams data with the injury_data on the team name
games_missed_from_injury <- merge(games_missed_from_injury, nfl_teams, by.x = "Team", by.y = "team_name")


## creation of unique id: 'player_id'
games_missed_from_injury$player_id <- paste0(games_missed_from_injury$season, "_", games_missed_from_injury$last_name, "_", games_missed_from_injury$team_abbreviation)

#games_missed_from_injury = games_missed_from_injury %>% filter(season == 2019)


#write_csv(games_missed_from_injury, 'games_missed_from_injury_2019_2022_complete.csv')


```

```{r pulling_position_and_team_name_of_injured_player_who_missed_game}

library(nflfastR)
player_position = fast_scraper_roster(2019: 2022)

player_position = player_position[,c('season', 'team', 'position','depth_chart_position', 'jersey_number',
                                     'status', 'full_name', 'first_name', 'last_name', 'birth_date',
                                     'college', 'years_exp', 'week', 'game_type', 'football_name', 'entry_year')]

colnames(player_position) = c('season', 'team', 'position','depth_chart_position', 'jersey_number',
                                     'status', 'full_name', 'first_name', 'last_name', 'birth_date',
                                     'college', 'years_exp', 'week', 'game_type', 'football_name', 'entry_year')

## creation of unique id: 'player_id' for the player_position table
player_position$player_id <- paste0(player_position$season, "_", player_position$last_name, "_", player_position$team)
player_position <- distinct(player_position, player_id, .keep_all = TRUE)
#write_csv(player_position, 'Rmd_files/player_position_Rmd1.csv')
## had to make manual changes to certain players names (i.e., Vander Esch from one of the tables, Esch from another; DeValve in one table was Devalve in another)

```



```{r merging_player_position_with_games_missed_from_injury}

library(dplyr)
## changed to 2019

games_missed_from_injury = read_csv('Independent Study/final_files/games_missed_from_injury_2019_2022_complete_update.csv')
names(games_missed_from_injury)[8] <- "team"

# Create a new column 'area_of_injury' as NA
games_missed_from_injury$area_of_injury <- NA

# Extract the area of injury from the 'Notes' column and assign it to the 'area_of_injury' column
games_missed_from_injury$area_of_injury <- ifelse(is.na(games_missed_from_injury$area_of_injury),
                                                  str_extract(games_missed_from_injury$Notes, ".*(?= injury)"),
                                                  games_missed_from_injury$area_of_injury)

# Replace remaining NA values with original values from 'Notes'
games_missed_from_injury$area_of_injury <- ifelse(is.na(games_missed_from_injury$area_of_injury),
                                                  games_missed_from_injury$Notes,
                                                  games_missed_from_injury$area_of_injury)

# Remove parenthesis from the 'area_of_injury' column
games_missed_from_injury$area_of_injury <- gsub("\\s*\\([^\\)]+\\)", "", games_missed_from_injury$area_of_injury)


## table for the frequency of certain types of injuries that resulted from players missing games
injury_counts <- table(games_missed_from_injury$area_of_injury)
injury_counts_sorted <- sort(injury_counts, decreasing = TRUE)


# Group the injury data by player_id, season, team, and any other relevant columns to see the number of games missed in a season by a player
games_missed_from_injury_agg <- games_missed_from_injury %>%
  group_by(player_id, season, team) %>%
  summarize(games_missed = n())
head(games_missed_from_injury_agg,20)

## updated csv file from manual changes to player names to match the 
player_position1 = read_csv('Independent Study/final_files/player_position_Rmd1.csv')
player_position1 <- distinct(player_position, player_id, .keep_all = TRUE)


# Merge the two tables by player ID, position, and team
merged_table_injury_by_position <- inner_join(games_missed_from_injury, player_position1, by = "player_id")
merged_table_injury_by_position <- merged_table_injury_by_position %>%
  select(-ends_with(".y")) %>%  # Remove the ".y" variables
  rename_with(~gsub("\\.x$", "", .), ends_with(".x"))  # Rename the ".x" variables

merged_table_injury_by_position <- merged_table_injury_by_position %>%
  select(-player_id, player_id)
head(merged_table_injury_by_position,20)


merged_table_num_injuries_by_player_by_season = inner_join(games_missed_from_injury_agg, player_position1, by = 'player_id')
head(merged_table_num_injuries_by_player_by_season,20)




# Merge the aggregated injury data with the player position data
#player_data <- left_join(games_missed_from_injury_agg,player_position ,by = "player_id")
## provides the number of injuries that a player suffered associated with the player_id of the individual
## results in 1423 observations, 44 missing observations, total games missed in the table: 2621

#player_data <- dplyr::full_join(games_missed_from_injury_agg,player_position ,by = "player_id")

#player_data <- merge(games_missed_from_injury_agg,player_position ,by = "player_id")
## results in 1378 observations, NO missing observations, total games missed in the table: 2566


#merged_data <- left_join(games_missed_from_injury, player_position, by = "player_id")
## results in 55 missing obs, 2621 observations

#colSums(is.na(merged_table_injury_by_position))
#write_csv(merged_table_injury_by_position, "Rmd_files/merged_table_injury_by_position1.csv")


#head(merged_table_injury_by_position,20)


#colSums(is.na(merged_table_num_injuries_by_player_by_season))
#write_csv(merged_table_num_injuries_by_player_by_season, "Rmd_files/merged_table_num_injuries_by_player_by_season1.csv")
#head(merged_table_injury_by_position,20)
```


```{r test_position}
df2 = read_csv("Independent Study/final_files/df_master_schedule_injury_surface_2019_23_weather_distance_days_since_last_game.csv")
## still need to add information about neutral sites along with correcting superbowl locations and associated information

## these changes have been made previously
#df2$Away_abbr <- ifelse(df2$Away_abbr == "LA", "LAR", df2$Away_abbr)
#df2$Home_abbr <- ifelse(df2$Home_abbr == "LA", "LAR", df2$Home_abbr)
#df2$game_id <- gsub("LA", "LAR", df2$game_id)
df2_players_sep = df2 %>% separate_rows(injured_players, sep = ",\\s*")
#write_csv(df2_players_sep, "Rmd_files/changing_suffixes2.csv")


df2_players_sep_update1 = read_csv("Independent Study/final_files/changing_suffixes2_update.csv")
## made corrections to names that contained suffixes and those that had periods in them (i.e., St. Brown)

df2_players_sep_update1 <- df2_players_sep_update1 %>%
  separate(col = injured_players, into = c('injured_first_name', 'injured_last_name'), sep = '\\.') %>%
  mutate(injured_last_name = if_else(str_detect(injured_last_name, '\\s'), 
                                     str_split(injured_last_name, '\\s')[[1]][1], 
                                     injured_last_name))

df2_players_sep_update1$injured_last_name <- ifelse(df2_players_sep_update1$injured_last_name == "NA", "No Injury", df2_players_sep_update1$injured_last_name)
df2_players_sep_update1$injured_last_name[is.na(df2_players_sep_update1$injured_last_name)] <- "No Injury"
## next code is to add the player position to contact_noncontact and then add them all back to df2_players_sep

## need to make sure that the abbreviations and game_id later used to merge are the same for both datasets
df2_players_sep_update1$game_id <- gsub("OAK", "LV", df2_players_sep_update1$game_id)
df2_players_sep_update1$Away_abbr <- ifelse(df2_players_sep_update1$Away_abbr == "OAK", "LV", df2_players_sep_update1$Away_abbr)
df2_players_sep_update1$Home_abbr <- ifelse(df2_players_sep_update1$Home_abbr == "OAK", "LV", df2_players_sep_update1$Home_abbr)

any(duplicated(df2_players_sep_update1))
df2_updated = distinct(df2_players_sep_update1)



#df2_players_sep = df2_players_sep %>% 
  #tidyr::separate(col = injured_team, into = c('injured_team', 'injured_player_num'))
injury_test2 = read_csv("Independent Study/final_files/contact_noncontact_2019_2020_full_seasons.csv")
#injury_test2 = read_csv("injury_test_neutral_sites.csv")
#error_testing = read_csv("Rmd_files/error_checking.csv")

injury_test2 = injury_test2 %>% 
  tidyr::separate(col = injured_team, into = c('injured_team', 'injured_player_num'))


#write_csv(injury_test2, "Rmd_files/joiner2_contact_corrections.csv")
injury_test2_update = read_csv("Independent Study/final_files/joiner2_contact_corrections_update.csv")

injury_test2_update$game_id <- gsub("OAK", "LV", injury_test2_update$game_id)
injury_test2_update$injured_team <- ifelse(injury_test2_update$injured_team == "OAK", "LV", injury_test2_update$injured_team)
injury_test2_update$home_team <- ifelse(injury_test2_update$home_team == "OAK", "LV", injury_test2_update$home_team)
injury_test2_update$away_team <- ifelse(injury_test2_update$away_team == "OAK", "LV", injury_test2_update$away_team)

injury_test2_update <- injury_test2_update %>%
  separate(col = injured_player, into = c('injured_first_name', 'injured_last_name'), sep = '\\.') %>%
  mutate(injured_last_name = if_else(str_detect(injured_last_name, '\\s'), 
                                     str_split(injured_last_name, '\\s')[[1]][1], 
                                     injured_last_name))

#write_csv(injury_test2_update, "changing_game_id_injury_test2.csv")

#injury_test2_update1 = read_csv("changing_game_id_injury_test2.csv")

#injury_test2 <- injury_test2 %>% 
  #separate(col = injured_player, into = c('injured_first_name', 'injured_last_name'), sep = '\\.(?=\\S+$)', extra = 'merge') %>% 
  #mutate(injured_last_name = ifelse(grepl("[-']", injured_last_name), paste(injured_last_name, sep = ""), injured_last_name))
#injury_test2 = injury_test2[,-29]

#write_csv(injury_test2, "test_injury_merge1.csv")
## changing names that contain II, III, Jr. Sr. etc
#injury_update = read_csv("test_injury_merge1.csv")
## changing LAR to LA in the game_id


injury_test2_update$season = substr(injury_test2_update$game_id, start = 1, stop = 4)
injury_test2_update$player_id = paste0(injury_test2_update$season,"_", injury_test2_update$injured_team, "_", injury_test2_update$injured_player_num, "_", injury_test2_update$injured_last_name)

#injury_test2_update$player_id = paste0(injury_test2_update$season,"_", injury_test2_update$injured_team, "_", injury_test2_update$injured_player_num)


rosters = nflreadr::load_rosters(2019:2022)
rosters$team <- ifelse(rosters$team == "LA", "LAR", rosters$team)
rosters = rosters %>% filter(season %in% c(2019,2020))
#rosters$team <- ifelse(rosters$team == "OAK", "LV", rosters$team)
rosters$player_id = paste0(rosters$season, "_", rosters$team, "_", rosters$jersey_number, "_", rosters$last_name)
#rosters$player_id = paste0(rosters$season, "_", rosters$team, "_", rosters$jersey_number)


injury_test2_sorted <- injury_test2_update %>% arrange(player_id)
rosters_sorted <- rosters %>% arrange(player_id)

position_injury_player <- merge(injury_test2_update, rosters, by = "player_id")
colSums(is.na(position_injury_player))
rows_dropped <- anti_join(injury_test2_update, rosters, by = "player_id")
## currently droppoing 2 observations --> will fix later

## dropping variables that have excessive missing values or are unimportant to the analysis
position_injury_player = position_injury_player %>% 
  select(-c("gsis_id", "espn_id", "sportradar_id", "yahoo_id", "rotowire_id", "pff_id", "pfr_id",
                                                    "fantasy_data_id", "sleeper_id", "headshot_url", "ngs_position", "gsis_it_id", "status_description_abbr", "esb_id", "rookie_year", "draft_club", "draft_number", "week.y"))

unique(position_injury_player$injured_team)
#write_csv(position_injury_player, "position_injury_player_2019_2020_fixed.csv")



##joining position_injury_player with df2_players_sep to get the game information
pos_inj_player = read_csv("Independent Study/final_files/position_injury_player_2019_2020_fixed_update.csv")
pos_inj_player$connector = paste0(pos_inj_player$game_id, "_", pos_inj_player$injured_last_name)

library(lubridate)
#pos_inj_player$game_date = as.Date(pos_inj_player$game_date, origin = "1899-12-30")
pos_inj_player$date = format(as.Date(pos_inj_player$game_date,origin = "1899-12-31"), "%m/%d/%y")

df2_updated$connector = paste0(df2_updated$game_id, "_", df2_updated$injured_last_name)


game_injury_player = merge(pos_inj_player,df2_updated,  by = "connector")
colSums(is.na(game_injury_player))
rows_dropped <- anti_join(pos_inj_player, df2_updated, by = "connector")
injury_teams1 = distinct(game_injury_player, injured_team)
colSums(is.na(game_injury_player))

#write_csv(game_injury_player, "game_injury_player_2019_2020_complete_final.csv")

```


```{r spreads_and_totals_combined}
library(nflfastR)
vegas_odds <- nflfastR::load_pbp(2019:2022)
vegas_odds = vegas_odds %>% select(game_id,home_team,away_team,season_type,week,quarter_seconds_remaining,
                     half_seconds_remaining,game_seconds_remaining,game_half,qtr,down,ydstogo, yrdln,spread_line, total_line, side_of_field, defteam,season)

vegas_odds$home_team <- ifelse(vegas_odds$home_team == "LA", "LAR", vegas_odds$home_team)
vegas_odds$away_team <- ifelse(vegas_odds$away_team == "LA", "LAR", vegas_odds$away_team)

# Replace "LA" with "LAR" in game_id
#vegas_odds2 <- vegas_odds %>%
  #mutate(game_id = str_replace(game_id, "(?<=^\\d{4}_\\d{2}_)[^_]*\\bLA\\b[^_]*", "LAR"))

#vegas_odds

#write_csv(vegas_odds, "vegas_odds_corrections2.csv")

vegas_update = read_csv("Independent Study/final_files/vegas_odds_corrections2_update.csv")
vegas_update = distinct(vegas_update)

## changed numeric weeks like 18 (in 2019 and 2022) and 19 as WildCard, Division, etc.
game_injury_no_odds1 = read_csv("Independent Study/final_files/game_injury_player_2019_2020_complete_final_update.csv")
game_injury_no_odds1 = game_injury_no_odds1 %>% 
  select(-c("week...48", "season...60"))
game_injury_no_odds1 = rename(game_injury_no_odds1, week = week...7)
game_injury_no_odds1 = rename(game_injury_no_odds1, season = season...30)
game_injury_no_odds1 = distinct(game_injury_no_odds1)

#sum(!complete.cases(game_injury_no_odds1[, c("game_id", "week", "quarter_seconds_remaining", "half_seconds_remaining", "game_seconds_remaining","yrdln","season_type", "game_half", "home_team", "away_team")]))

#sum(!complete.cases(vegas_update[, c("game_id", "week", "quarter_seconds_remaining", "half_seconds_remaining", "game_seconds_remaining","yrdln","season_type", "game_half", "home_team", "away_team")]))

sum(duplicated(game_injury_no_odds1))

temp_merge1= inner_join(game_injury_no_odds1, vegas_update, by = c("game_id", "week", "quarter_seconds_remaining", "half_seconds_remaining", "game_seconds_remaining","yrdln","season_type", "game_half", "home_team", "away_team"))
## 30 observations dropped (1.9% of total dataset)
temp_merge1 = distinct(temp_merge1)



#write_csv(temp_merge1, "Rmd_files/game_injury_player_2019_2020_complete_update_vegas.csv")
rows_dropped1 =  anti_join(game_injury_no_odds1, vegas_update, by = c("game_id", "week", "quarter_seconds_remaining", "half_seconds_remaining", "game_seconds_remaining","yrdln","season_type", "game_half", "home_team", "away_team")) %>% as.data.frame()
sum(duplicated(rows_dropped1))
rows_dropped1 = distinct(rows_dropped1)


#rows_dropped2$home_team <- gsub("LV", "OAK",rows_dropped2$home_team )
#rows_dropped2$away_team <- gsub("LV", "OAK",rows_dropped2$away_team )
#write_csv(rows_dropped2, "rows_dropped_update_LV_to_OAK2.csv")

#dropped_merge = read_csv("rows_dropped_update_LV_to_OAK2_update.csv")

#temp_merge3 = inner_join(dropped_merge, vegas_update, by = c("game_id", "week", "quarter_seconds_remaining", "half_seconds_remaining", "game_seconds_remaining","season_type","yrdln" ,"game_half", "home_team", "away_team"))

#temp_merge1 = temp_merge1 %>% 
  #select(-c("season...60"))
#temp_merge3 = temp_merge3 %>% 
  #select(-c("season...59"))

#temp_merge1 = rename(temp_merge1, season = season...30)
#temp_merge3 = rename(temp_merge3, season = season...30)


#complete_df = rbind(temp_merge1,temp_merge3)

#write_csv(temp_merge1, "game_injury_player_2019_2020_complete_update_vegas.csv")


```

```{r weeks_bucketed}

game_injury_odds = read_csv("Independent Study/final_files/game_injury_player_2019_2020_complete_update_vegas_update.csv")

# use case_when to bucket weeks into new variable weeks_bucketed
game_injury_odds$weeks_bucketed <- as.character(case_when(
  game_injury_odds$week %in% 1:4  ~ 1,
  game_injury_odds$week %in% 5:8  ~ 2,
  game_injury_odds$week %in% 9:12 ~ 3,
  game_injury_odds$week %in% 13:17 ~ 4,
  game_injury_odds$week %in% 18:21 ~ 5
)
)

# view the results
unique(game_injury_odds$weeks_bucketed)

#write_csv(game_injury_odds, "game_injury_player_2019_2020_complete_update_vegas_weeks_bucketed.csv")

```


